Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-14 WSSSPE1_report/WSSSPE1_summary_report.tex
make sure citations don't break lines

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as
\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

\item Mark C. Miller, Lori Diachin, Satish Balay, Lois Curfman
  McInnes, Barry Smith. Package Management Practices Essential for
  Interoperability: Lessons Learned and Strategies Developed for
  FASTMath \cite{Miller_WSSSPE}

\item Karl W. Broman, Thirteen years of R/qtl: Just barely sustainable
  \cite{Broman_WSSSPE}

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures \cite{Ferenbaugh_WSSSPE}

\item Eric G Stephan, Todd O Elsethagen, Kerstin Kleese van Dam, Laura
  Riihimaki. What Comes First, the OWL or the Bean?
  \cite{Stephan_WSSSPE}

\item Derek R. Gaston, John Peterson, Cody J. Permann, David Andrs,
  Andrew E. Slaughter, Jason M. Miller, Continuous Integration for
  Concurrent Computational Framework and Application Development
  \cite{Gaston_WSSSPE}

\item Anshu Dubey, B. Van Straalen. Experiences from Software
  Engineering of Large Scale AMR Multiphysics Code Frameworks
  \cite{Dubey_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development \cite{Blatt_WSSSPE}

@@

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software \cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software
  \cite{Choi_WSSSPE}

\item Michael Crusoe, C. Titus Brown. Walking the talk: adopting and
  adapting sustainable scientific software development processes in a
  small biology lab \cite{Crusoe_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava Majumdar.
The MVAPICH Project: Evolution and Sustainability of an Open Source
Production Quality MPI Library for HPC \cite{Panda_WSSSPE}

\item Eric M. Heien, Todd L. Miller, Becky Gietzel, Louise
  H. Kellogg. Experiences with Automated Build and Test for
  Geodynamics Simulation Codes \cite{Heien_WSSSPE}

@@

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software \cite{Trainer_WSSSPE}

\item Jeremy Cohen, Chris Cantwell, Neil Chue Hong, David Moxey,
  Malcolm Illingworth, Andrew Turner, John Darlington, Spencer
  Sherwin. Simplifying the Development, Use and Sustainability of HPC
  Software \cite{Cohen_WSSSPE}

\item Jaroslaw Slawinski, Vaidy Sunderam. Towards Semi-Automatic
  Deployment of Scientific and Engineering Applications
  \cite{Slawinski_WSSSPE}

@@

\item Andreas Prli\'{c}, James B. Procter. Ten Simple Rules for the
  Open Development of Scientific Software \cite{Prlic_WSSSPE}

\item Anshu Dubey, S. Brandt, R. Brower, M. Giles, P. Hovland,
  D. Q. Lamb, F. Löffler, B. Norris, B. O'Shea, C. Rebbi, M. Snir,
  R. Thakur, Software Abstractions and Methodologies for HPC
  Simulation Codes on Future Architectures \cite{Dubey2_WSSSPE}

\item Jeffrey Carver, George K. Thiruvathukal. Software Engineering
  Need Not Be Difficult \cite{Carver_WSSSPE}

\item Craig A. Stewart, Julie Wernert, Eric A. Wernert, William
  K. Barnett, Von Welch. Initial Findings from a Study of Best
  Practices and Models for Cyberinfrastructure Software Sustainability
  \cite{Stewart_WSSSPE}

\item Jed Brown, Matthew Knepley, Barry Smith. Run-time extensibility:
  anything less is unsustainable \cite{Brown_WSSSPE}

\item Shel Swenson, Yogesh Simmhan, Viktor Prasanna, Manish Parashar,
  Jason Riedy, David Bader, Richard Vuduc. Sustainable Software
  Development for Next-Gen Sequencing (NGS) Bioinformatics on Emerging
  Platforms \cite{Swenson_WSSSPE}

@@

\item Coral Calero, M. Angeles Moraga, Manuel F. Bertoa. Towards a
  Software Product Sustainability Model \cite{Calero_WSSSPE}

\item Colin C. Venters, Lydia Lau, Michael K. Griffiths, Violeta
  Holmes, Rupert R. Ward, Jie Xu. The Blind Men and the Elephant:
  Towards a Software Sustainability Architectural Evaluation Framework
  \cite{Venters_WSSSPE}

\item Marlon Pierce, Suresh Marru, Chris Mattmann. Sustainable
  Cyberinfrastructure Software Through Open Governance
  \cite{Pierce_WSSSPE}

\item Daniel S. Katz, David Proctor. A Framework for Discussing
  e-Research Infrastructure Sustainability \cite{Katz_WSSSPE}

\item Christopher Lenhardt, Stanley Ahalt, Brian Blanton, Laura
  Christopherson, Ray Idaszak. Data Management Lifecycle and Software
  Lifecycle Management in the Context of Conducting Science
  \cite{Lenhardt_WSSSPE}

\item Nicholas Weber, Andrea Thomer, Michael Twidale. Niche Modeling:
  Ecological Metaphors for Sustainable Software in Science
  \cite{Weber_WSSSPE}

@@

The policy panel and discussion addressed the question of recognition of work on scientific software and linked that recognition to questions of reward and thus motivation for particular kinds of work on scientific software. In short, software work in science was seen to be inadequately visible in ways that ``count'' within the reputation system underlying science. In his paper for this workshop, Katz placed software work along with other ``activities that facilitate science but are not currently rewarded or recognized'' \cite{Katz_WSSSPE}. Priem and Piwowar argued for the need to ``support all researchers in presenting meaningful impact evidence in tenure, promotion, and funding applications.'' \cite{Priem_WSSSPE}.  Kneply et al. argued that the visibility of software that supported a piece of science ``can have detrimental effects on funding, future library development, and even scientific careers.'' \cite{Knepley_WSSSPE}.

These papers, and the discussion at the workshop, join a nascent literature seeking to understand what drives software work in science and how the reward systems of science thereby shape the type of software work undertaken, including the extent to which developers are motivated to build software for the long-term, for the use of others and whether to work collaboratively or separately \cite{howison_incentives_2013, howison_scientific_2011, bietz_synergizing_2010}. Software work is not only motivated by direct citations, but the visibility of software work in the literature is important to those who write software used in science.

@@

Most of the papers in this area focused on visibility of software in scientific papers, since scientific papers are the most widely accepted documentation of achievement in science. It was noted that there appear to be no widely accepted standards on how the use of software towards a paper ought to be mentioned, that journals, citation style guides and other guides to scientific conduct are vague about how to describe software. To address this, papers spoke of a need for a fixed identifier for software, either directly through a mechanism such as a Digital Object Identifier \cite{Katz_WSSSPE,Knepley_WSSSPE} or via a published paper written to document the software (and perhaps its creation), a ``software paper'' \cite{Chue_Hong_WSSSPE}.

Another approach is to try to reduce the difficulty of citing software for authors, acknowledging that authors are often working with software that itself wraps other software and therefore hides that software. Knepley et al approach this by proposing a mechanism by which the software itself, after it has run, provides the user with a set of citations that are relevant to the code actually run. They describe a prototype implementation whereby the citations are embedded in libraries and reported along with the results, via a commandline interface \cite{Knepley_WSSSPE}. Discussion highlighted the difficulty that attempting to acknowledge the contributions of all pieces of dependent code within a paper faces the difficulty of creating very long citation lists, straining the analogy of code used to papers cited. Katz approaches this issue by proposing a system of ``transitive credit,'' recording dependencies and relative contributions outside particular papers, relieving authors from the responsibility of acknowledging each and every dependency. In stead authors would acknowledge the percentage contribution of the software they used directly and an external system would then be able to recursively allocate that credit to those who had provided dependencies. Finally Priem and Piwowar argued that machine learning techniques could examine the body of published literature and extract mentions of software, coping with the multitude of informal ways in which authors mention software they have used \cite{Priem_WSSSPE}.  Discussion included turning the emphasis from asking users to improve their citation of software contributions, to ask how projects producing software might monitor the literature to improve their ability to show impact. Michael McLennan described the approach taken by the NanoHub project to scan the literature using keywords and names of known users to discover papers that are likely to have used their software and platform, then to have graduate students read each paper, highlighting any mention and perhaps following up with the authors to locate stories of impact.  Work since the workshop has described this process at publications.wikia.com.

@@

Priem and Piwowar argued for additionally looking beyond publications and drawing on evidence of contribution, and impact, recorded in other online systems, such as source code repositories like github, both code contributions, downloads of code and dependencies as well as conversations about software (mailing lists, twitter and beyond) \cite{Priem_WSSSPE}. They argue for providing scholars with flexible resources so that they can tell their own stories in the manner most appropriate for them and their audiences, a principle of the ``alt.metrics'' approach.

@@

\item Matthew Knepley, Jed Brown, Lois Curfman McInnes, Barry
  Smith. Accurately Citing Software and Algorithms Used in
  Publications \cite{Knepley_WSSSPE}

\item Jason Priem, Heather Piwowar. Toward a comprehensive impact
  report for every software project \cite{Priem_WSSSPE}

\item Daniel S. Katz. Citation and Attribution of Digital Products:
  Social and Technological Concerns \cite{Katz2_WSSSPE}

\item Neil Chue Hong, Brian Hole, Samuel Moore. Software Papers:
  improving the reusability and sustainability of scientific software
  \cite{Chue_Hong_WSSSPE}

@@

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software
  \cite{Loffler_WSSSPE}

@@

\item Victoria Stodden, Sheila Miguez. Best Practices for
  Computational Science: Software Infrastructure and Environments for
  Reproducible and Extensible Research \cite{Stodden_WSSSPE}

@@

\item Randy Heiland, Betsy Thomas, Von Welch, Craig Jackson. Toward a
  Research Software Security Maturity Model \cite{Heiland_WSSSPE}

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software \cite{Blanton_WSSSPE}

\item Daisie Huang, Hilmar Lapp. Software Engineering as
  Instrumentation for the Long Tail of Scientific Software
  \cite{Huang_WSSSPE}

\item Rich Wolski, Chandra Krintz, Hiranya Jayathilaka, Stratos
  Dimopoulos, Alexander Pucher. Developing Systems for API Governance
  \cite{Wolski_WSSSPE}

@@

Drawing on experiences from high-energy physics, \cite{Vay_WSSSPE} proposed
developing teams of technical specialists able to overcome a lack of
coordination between projects.  Maximization of scientific output requires
maximizing the usability of the scientific codes while minimizing the cost of
developing and supporting those codes.  This included targeting different
architectures for their software to be deployed, as well as coordination
between technically-focused individuals and usage of a common scripting
language between projects.  Instead of fragmenting the development of
simulation codes across institutions, the paper suggests that a cohesive
strategy reducing duplication and increasing coordination will broadly increase
the efficiency across institutions.  The approach proposed is of de-fragmenting
the existing ecosystem in a non-disruptive way.

@@

In \cite{Hanwell_WSSSPE}, the authors reflect on the 15-year history of open
source software development at Kitware.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions.  This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables Kitware to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

At NESCent, a combination of in-house informatics individuals and domain
scientists collaborate to develop software to study evolutionary science.  The
report, \cite{Cranston_WSSSPE}, studied the success of a ``hackathon'' model
for development, where short-form, hands-on events combining users,
researcher-developers and software engineers targeted specific code
improvements.  From this experiment, the authors identified sweveral key
outcomes as well as lessons-learned -- specifically, the co-localization of
developers was seen as having a strong impact, enabling casual conversation
that led to discrete outcomes.  The formation of the discussion mailing list,
specifically in response to the social capital built at the hackathon, was seen
as spurring on longer-term benefits to the community and fostering
sustainability.

@@

\item Reagan Moore. Extensible Generic Data Management Software
  \cite{Moore_WSSSPE}

\item Karen Cranston, Todd Vision, Brian O'Meara, Hilmar Lapp. A
  grassroots approach to software sustainability
  \cite{Cranston_WSSSPE}

\item J.-L. Vay, C. G. R. Geddes, A. Koniges, A. Friedman,
  D. P. Grote, D. L. Bruhwiler. White Paper on DOE-HEP Accelerator
  Modeling Science Activities \cite{Vay_WSSSPE}

\item Ketan Maheshwari, David Kelly, Scott J. Krieder, Justin M. Wozniak, Daniel S. Katz, Mei Zhi-Gang, Mainak Mookherjee. Reusability in Science: From Initial User Engagement to Dissemination of Results \cite{Maheshwari_WSSSPE}

\item Edmund Hart, Carl Boettiger, Karthik Ram, Scott Chamberlain. rOpenSci -- a collaborative effort to develop R-based tools for facilitating Open Science \cite{Hart_WSSSPE}

@@

\item L. Christopherson, R. Idaszak, S. Ahalt. Developing Scientific Software through the Open Community Engagement Process \cite{Christopherson_WSSSPE}

\item Marlon Pierce, Suresh Marru, Mark A. Miller, Amit Majumdar, Borries Demeler. Science Gateway Operational Sustainability: Adopting a Platform-as-a-Service Approach \cite{Pierce2_WSSSPE}

\item Lynn Zentner, Michael Zentner, Victoria Farnsworth, Michael
  McLennan, Krishna Madhavan, and Gerhard Klimeck, nanoHUB.org:
  Experiences and Challenges in Software Sustainability for a Large
  Scientific Community \cite{Zentner_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community
  \cite{Terrel_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software
  \cite{Loffler_WSSSPE}

\item Nancy Wilkins-Diehr, Katherine Lawrence, Linda Hayden, Marlon Pierce, Suresh Marru, Michael McLennan, Michael Zentner, Rion Dooley, Dan Stanzione. Science Gateways and the Importance of Sustainability \cite{Wilkins-Diehr_WSSSPE}

@@

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science \cite{Hanwell_WSSSPE}

@@

\item Anne C. Elster. Software for Science: Some Personal Reflections
  \cite{Elster_WSSSPE}

\item Ian Foster, Vas Vasiliadis, Steven Tuecke. Software as a Service
  as a path to software sustainability \cite{Foster_WSSSPE}

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science \cite{Hanwell_WSSSPE}

@@

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software \cite{Blanton_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development \cite{Blatt_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava
  Majumdar. The MVAPICH Project: Evolution and Sustainability of an
  Open Source Production Quality MPI Library for HPC
  \cite{Panda_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community
  \cite{Terrel_WSSSPE}

@@

\item Ivan Girotto, Axel Kohlmeyer, David Grellscheid, Shawn
  T. Brown. Advanced Techniques for Scientific Programming and
  Collaborative Development of Open Source Software Packages at the
  International Centre for Theoretical Physics (ICTP)
  \cite{Girotto_WSSSPE}

\item Thomas Crawford. On the Development of Sustainable Software for
  Computational Chemistry \cite{Crawford_WSSSPE}

@@

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures \cite{Ferenbaugh_WSSSPE}

@@

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software \cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software
  \cite{Choi_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software
  \cite{Loffler_WSSSPE}

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software \cite{Trainer_WSSSPE}

>>>>

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as~\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

\item Mark C. Miller, Lori Diachin, Satish Balay, Lois Curfman
  McInnes, Barry Smith. Package Management Practices Essential for
  Interoperability: Lessons Learned and Strategies Developed for
  FASTMath~\cite{Miller_WSSSPE}

\item Karl W. Broman, Thirteen years of R/qtl: Just barely sustainable~\cite{Broman_WSSSPE}

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures~\cite{Ferenbaugh_WSSSPE}

\item Eric G Stephan, Todd O Elsethagen, Kerstin Kleese van Dam, Laura
  Riihimaki. What Comes First, the OWL or the Bean?~\cite{Stephan_WSSSPE}

\item Derek R. Gaston, John Peterson, Cody J. Permann, David Andrs,
  Andrew E. Slaughter, Jason M. Miller, Continuous Integration for
  Concurrent Computational Framework and Application Development~\cite{Gaston_WSSSPE}

\item Anshu Dubey, B. Van Straalen. Experiences from Software
  Engineering of Large Scale AMR Multiphysics Code Frameworks~\cite{Dubey_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development~\cite{Blatt_WSSSPE}

@@

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software~\cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\item Michael Crusoe, C. Titus Brown. Walking the talk: adopting and
  adapting sustainable scientific software development processes in a
  small biology lab~\cite{Crusoe_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava Majumdar.
The MVAPICH Project: Evolution and Sustainability of an Open Source
Production Quality MPI Library for HPC~\cite{Panda_WSSSPE}

\item Eric M. Heien, Todd L. Miller, Becky Gietzel, Louise
  H. Kellogg. Experiences with Automated Build and Test for
  Geodynamics Simulation Codes~\cite{Heien_WSSSPE}

@@

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software~\cite{Trainer_WSSSPE}

\item Jeremy Cohen, Chris Cantwell, Neil Chue Hong, David Moxey,
  Malcolm Illingworth, Andrew Turner, John Darlington, Spencer
  Sherwin. Simplifying the Development, Use and Sustainability of HPC
  Software~\cite{Cohen_WSSSPE}

\item Jaroslaw Slawinski, Vaidy Sunderam. Towards Semi-Automatic
  Deployment of Scientific and Engineering Applications~\cite{Slawinski_WSSSPE}

@@

\item Andreas Prli\'{c}, James B. Procter. Ten Simple Rules for the
  Open Development of Scientific Software~\cite{Prlic_WSSSPE}

\item Anshu Dubey, S. Brandt, R. Brower, M. Giles, P. Hovland,
  D. Q. Lamb, F. Löffler, B. Norris, B. O'Shea, C. Rebbi, M. Snir,
  R. Thakur, Software Abstractions and Methodologies for HPC
  Simulation Codes on Future Architectures~\cite{Dubey2_WSSSPE}

\item Jeffrey Carver, George K. Thiruvathukal. Software Engineering
  Need Not Be Difficult~\cite{Carver_WSSSPE}

\item Craig A. Stewart, Julie Wernert, Eric A. Wernert, William
  K. Barnett, Von Welch. Initial Findings from a Study of Best
  Practices and Models for Cyberinfrastructure Software Sustainability~\cite{Stewart_WSSSPE}

\item Jed Brown, Matthew Knepley, Barry Smith. Run-time extensibility:
  anything less is unsustainable~\cite{Brown_WSSSPE}

\item Shel Swenson, Yogesh Simmhan, Viktor Prasanna, Manish Parashar,
  Jason Riedy, David Bader, Richard Vuduc. Sustainable Software
  Development for Next-Gen Sequencing (NGS) Bioinformatics on Emerging
  Platforms~\cite{Swenson_WSSSPE}

@@

\item Coral Calero, M. Angeles Moraga, Manuel F. Bertoa. Towards a
  Software Product Sustainability Model~\cite{Calero_WSSSPE}

\item Colin C. Venters, Lydia Lau, Michael K. Griffiths, Violeta
  Holmes, Rupert R. Ward, Jie Xu. The Blind Men and the Elephant:
  Towards a Software Sustainability Architectural Evaluation Framework~\cite{Venters_WSSSPE}

\item Marlon Pierce, Suresh Marru, Chris Mattmann. Sustainable
  Cyberinfrastructure Software Through Open Governance~\cite{Pierce_WSSSPE}

\item Daniel S. Katz, David Proctor. A Framework for Discussing
  e-Research Infrastructure Sustainability~\cite{Katz_WSSSPE}

\item Christopher Lenhardt, Stanley Ahalt, Brian Blanton, Laura
  Christopherson, Ray Idaszak. Data Management Lifecycle and Software
  Lifecycle Management in the Context of Conducting Science~\cite{Lenhardt_WSSSPE}

\item Nicholas Weber, Andrea Thomer, Michael Twidale. Niche Modeling:
  Ecological Metaphors for Sustainable Software in Science~\cite{Weber_WSSSPE}

@@

The policy panel and discussion addressed the question of recognition of work on scientific software and linked that recognition to questions of reward and thus motivation for particular kinds of work on scientific software. In short, software work in science was seen to be inadequately visible in ways that ``count'' within the reputation system underlying science. In his paper for this workshop, Katz placed software work along with other ``activities that facilitate science but are not currently rewarded or recognized''~\cite{Katz_WSSSPE}. Priem and Piwowar argued for the need to ``support all researchers in presenting meaningful impact evidence in tenure, promotion, and funding applications.''~\cite{Priem_WSSSPE}.  Kneply et al. argued that the visibility of software that supported a piece of science ``can have detrimental effects on funding, future library development, and even scientific careers.''~\cite{Knepley_WSSSPE}.

These papers, and the discussion at the workshop, join a nascent literature seeking to understand what drives software work in science and how the reward systems of science thereby shape the type of software work undertaken, including the extent to which developers are motivated to build software for the long-term, for the use of others and whether to work collaboratively or separately~\cite{howison_incentives_2013, howison_scientific_2011, bietz_synergizing_2010}. Software work is not only motivated by direct citations, but the visibility of software work in the literature is important to those who write software used in science.

@@

Most of the papers in this area focused on visibility of software in scientific papers, since scientific papers are the most widely accepted documentation of achievement in science. It was noted that there appear to be no widely accepted standards on how the use of software towards a paper ought to be mentioned, that journals, citation style guides and other guides to scientific conduct are vague about how to describe software. To address this, papers spoke of a need for a fixed identifier for software, either directly through a mechanism such as a Digital Object Identifier~\cite{Katz_WSSSPE,Knepley_WSSSPE} or via a published paper written to document the software (and perhaps its creation), a ``software paper''~\cite{Chue_Hong_WSSSPE}.

Another approach is to try to reduce the difficulty of citing software for authors, acknowledging that authors are often working with software that itself wraps other software and therefore hides that software. Knepley et al approach this by proposing a mechanism by which the software itself, after it has run, provides the user with a set of citations that are relevant to the code actually run. They describe a prototype implementation whereby the citations are embedded in libraries and reported along with the results, via a commandline interface~\cite{Knepley_WSSSPE}. Discussion highlighted the difficulty that attempting to acknowledge the contributions of all pieces of dependent code within a paper faces the difficulty of creating very long citation lists, straining the analogy of code used to papers cited. Katz approaches this issue by proposing a system of ``transitive credit,'' recording dependencies and relative contributions outside particular papers, relieving authors from the responsibility of acknowledging each and every dependency. In stead authors would acknowledge the percentage contribution of the software they used directly and an external system would then be able to recursively allocate that credit to those who had provided dependencies. Finally Priem and Piwowar argued that machine learning techniques could examine the body of published literature and extract mentions of software, coping with the multitude of informal ways in which authors mention software they have used~\cite{Priem_WSSSPE}.  Discussion included turning the emphasis from asking users to improve their citation of software contributions, to ask how projects producing software might monitor the literature to improve their ability to show impact. Michael McLennan described the approach taken by the NanoHub project to scan the literature using keywords and names of known users to discover papers that are likely to have used their software and platform, then to have graduate students read each paper, highlighting any mention and perhaps following up with the authors to locate stories of impact.  Work since the workshop has described this process at publications.wikia.com.

@@

Priem and Piwowar argued for additionally looking beyond publications and drawing on evidence of contribution, and impact, recorded in other online systems, such as source code repositories like github, both code contributions, downloads of code and dependencies as well as conversations about software (mailing lists, twitter and beyond)~\cite{Priem_WSSSPE}. They argue for providing scholars with flexible resources so that they can tell their own stories in the manner most appropriate for them and their audiences, a principle of the ``alt.metrics'' approach.

@@

\item Matthew Knepley, Jed Brown, Lois Curfman McInnes, Barry
  Smith. Accurately Citing Software and Algorithms Used in
  Publications~\cite{Knepley_WSSSPE}

\item Jason Priem, Heather Piwowar. Toward a comprehensive impact
  report for every software project~\cite{Priem_WSSSPE}

\item Daniel S. Katz. Citation and Attribution of Digital Products:
  Social and Technological Concerns~\cite{Katz2_WSSSPE}

\item Neil Chue Hong, Brian Hole, Samuel Moore. Software Papers:
  improving the reusability and sustainability of scientific software~\cite{Chue_Hong_WSSSPE}

@@

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

@@

\item Victoria Stodden, Sheila Miguez. Best Practices for
  Computational Science: Software Infrastructure and Environments for
  Reproducible and Extensible Research~\cite{Stodden_WSSSPE}

@@

\item Randy Heiland, Betsy Thomas, Von Welch, Craig Jackson. Toward a
  Research Software Security Maturity Model~\cite{Heiland_WSSSPE}

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software~\cite{Blanton_WSSSPE}

\item Daisie Huang, Hilmar Lapp. Software Engineering as
  Instrumentation for the Long Tail of Scientific Software~\cite{Huang_WSSSPE}

\item Rich Wolski, Chandra Krintz, Hiranya Jayathilaka, Stratos
  Dimopoulos, Alexander Pucher. Developing Systems for API Governance~\cite{Wolski_WSSSPE}

@@

Drawing on experiences from high-energy physics,~\cite{Vay_WSSSPE} proposed
developing teams of technical specialists able to overcome a lack of
coordination between projects.  Maximization of scientific output requires
maximizing the usability of the scientific codes while minimizing the cost of
developing and supporting those codes.  This included targeting different
architectures for their software to be deployed, as well as coordination
between technically-focused individuals and usage of a common scripting
language between projects.  Instead of fragmenting the development of
simulation codes across institutions, the paper suggests that a cohesive
strategy reducing duplication and increasing coordination will broadly increase
the efficiency across institutions.  The approach proposed is of de-fragmenting
the existing ecosystem in a non-disruptive way.

@@

In~\cite{Hanwell_WSSSPE}, the authors reflect on the 15-year history of open
source software development at Kitware.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions.  This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables Kitware to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

At NESCent, a combination of in-house informatics individuals and domain
scientists collaborate to develop software to study evolutionary science.  The
report~\cite{Cranston_WSSSPE} studied the success of a ``hackathon'' model
for development, where short-form, hands-on events combining users,
researcher-developers and software engineers targeted specific code
improvements.  From this experiment, the authors identified sweveral key
outcomes as well as lessons-learned -- specifically, the co-localization of
developers was seen as having a strong impact, enabling casual conversation
that led to discrete outcomes.  The formation of the discussion mailing list,
specifically in response to the social capital built at the hackathon, was seen
as spurring on longer-term benefits to the community and fostering
sustainability.

@@

\item Reagan Moore. Extensible Generic Data Management Software~\cite{Moore_WSSSPE}

\item Karen Cranston, Todd Vision, Brian O'Meara, Hilmar Lapp. A
  grassroots approach to software sustainability~\cite{Cranston_WSSSPE}

\item J.-L. Vay, C. G. R. Geddes, A. Koniges, A. Friedman,
  D. P. Grote, D. L. Bruhwiler. White Paper on DOE-HEP Accelerator
  Modeling Science Activities~\cite{Vay_WSSSPE}

\item Ketan Maheshwari, David Kelly, Scott J. Krieder, Justin M. Wozniak, Daniel S. Katz, Mei Zhi-Gang, Mainak Mookherjee. Reusability in Science: From Initial User Engagement to Dissemination of Results~\cite{Maheshwari_WSSSPE}

\item Edmund Hart, Carl Boettiger, Karthik Ram, Scott Chamberlain. rOpenSci -- a collaborative effort to develop R-based tools for facilitating Open Science~\cite{Hart_WSSSPE}

@@

\item L. Christopherson, R. Idaszak, S. Ahalt. Developing Scientific Software through the Open Community Engagement Process~\cite{Christopherson_WSSSPE}

\item Marlon Pierce, Suresh Marru, Mark A. Miller, Amit Majumdar, Borries Demeler. Science Gateway Operational Sustainability: Adopting a Platform-as-a-Service Approach~\cite{Pierce2_WSSSPE}

\item Lynn Zentner, Michael Zentner, Victoria Farnsworth, Michael
  McLennan, Krishna Madhavan, and Gerhard Klimeck, nanoHUB.org:
  Experiences and Challenges in Software Sustainability for a Large
  Scientific Community~\cite{Zentner_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community~\cite{Terrel_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

\item Nancy Wilkins-Diehr, Katherine Lawrence, Linda Hayden, Marlon Pierce, Suresh Marru, Michael McLennan, Michael Zentner, Rion Dooley, Dan Stanzione. Science Gateways and the Importance of Sustainability~\cite{Wilkins-Diehr_WSSSPE}

@@

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science~\cite{Hanwell_WSSSPE}

@@

\item Anne C. Elster. Software for Science: Some Personal Reflections~\cite{Elster_WSSSPE}

\item Ian Foster, Vas Vasiliadis, Steven Tuecke. Software as a Service
  as a path to software sustainability~\cite{Foster_WSSSPE}

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science~\cite{Hanwell_WSSSPE}

@@

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software~\cite{Blanton_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development~\cite{Blatt_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava
  Majumdar. The MVAPICH Project: Evolution and Sustainability of an
  Open Source Production Quality MPI Library for HPC~\cite{Panda_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community~\cite{Terrel_WSSSPE}

@@

\item Ivan Girotto, Axel Kohlmeyer, David Grellscheid, Shawn
  T. Brown. Advanced Techniques for Scientific Programming and
  Collaborative Development of Open Source Software Packages at the
  International Centre for Theoretical Physics (ICTP)~\cite{Girotto_WSSSPE}

\item Thomas Crawford. On the Development of Sustainable Software for
  Computational Chemistry~\cite{Crawford_WSSSPE}

@@

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures~\cite{Ferenbaugh_WSSSPE}

@@

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software~\cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software~\cite{Trainer_WSSSPE}
